---
title: "Attendance at Modelação Ecológica"
format: html
editor: visual
---

```{r,message=FALSE,warning=FALSE}
#load required packages
library(readxl)
library(tidyverse)
```

## Introduction

This document will track the attendance in ME course.

It illustrates a dynamic report, where each week data changes and hence the attendance changes too.

Data have been anonymized for RGPD reasons.

## Reading the data

Reading the data with function `read_excel` from package `readxl`:

```{r}
attend <- read_excel("TrackAttendance.xlsx")
attend <- cbind(1:nrow(attend),attend)
names(attend)[1]<-"student"
nclasses<-24
```

This year there will be `r nclasses` of Modelação Ecológica.

## Look at the data

Count how many students per class, and then plot that information

```{r}
#sum columns, except the first
index.not.aulas<-c(1,2)
plot(1:nclasses,colSums(attend[,-index.not.aulas]),ylab="Number of students",xlab="Class")
```

```{r file4readme,echo=FALSE,warning=FALSE,message=FALSE}
#not shown, but creates plot file for readme
jpeg(filename="file4readme.jpg")
#sum columns, except the first
index.not.aulas<-c(1,2)
plot(1:nclasses,colSums(attend[,-index.not.aulas]),ylab="Number of students",xlab="Class")
dev.off()
```

And we can see how the students are doing per MSc program. Or we might when we have enough data to do so!

```{r}
#placeholder for said analysis
```

# Modelling attendance

We can reshape the data from the "wide" to the "long" format, and use that to model the attendance as a function of the class. I also included the course for fun (but this can't be interpreted yet, as the records are not complete). Later we might want to include a student random effect, since presumably some students will have a higher tendency to show in class!

```{r}
attendtodate<-attend[,c(rep(TRUE,length(index.not.aulas)),colSums(attend[,-index.not.aulas])>0)]
attendtodatel<-gather(attendtodate,key=aula,value=present,aula1:tail(names(attendtodate),n=1))
```

and then implement the model

```{r}
glm1<-glm(present~aula+curso,data=attendtodatel,family=binomial)
summary(glm1)
```

Note some data cleaning is required before the analysis can be looked into. Some students were not registered in class and some classes were not delivered for reasons we do not control (e.g. Ciencias Research and Inovvation Day)